---
title: "Accuracy and Precision of Continuous Non-Invasive Arterial Pressure Monitoring Compared with Invasive or Non-Invasive Blood Pressure Monitoring"
subtitle: A systematic review and meta-analysis
bibliography: Extras/bibliography.bib
csl: Extras/sp-basic.csl
author:
  - Navpreet Kamboj MScN:
      institute: [LBFON]
      correspondence: true
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Kelly Metcalfe PhD:
      institute: [LBFON]
      correspondence: false
  - Charlene Chu PhD:
      institute: [LBFON]
      correspondence: false
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: false

institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia

output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
library(officer)
library(officedown)
library(drake)

loadd(nexfin_results)
loadd(nexfin_dbp_results)
loadd(nexfin_map_results)
loadd(primary_map_results)
loadd(primary_sbp_results)
loadd(primary_dbp_results)
loadd(finapres_map_results)
loadd(finapres_sbp_results)
loadd(finapres_dbp_results)
loadd(ncat_map_results)
loadd(ncat_sbp_results)
loadd(ncat_dbp_results)
loadd(tline_map_results)
loadd(tline_sbp_results)
loadd(tline_dbp_results)
loadd(cnap_map_results)
loadd(cnap_sbp_results)
loadd(cnap_dbp_results)
loadd(aat_map_results)
loadd(aat_sbp_results)
loadd(aat_dbp_results)
loadd(vc_map_results)
loadd(vc_sbp_results)
loadd(vc_dbp_results)
loadd(radial_map_results)
loadd(radial_sbp_results)
loadd(radial_dbp_results)
loadd(femoral_map_results)
loadd(femoral_sbp_results)
loadd(femoral_dbp_results)
```

\pagebreak


### Abstract

**Purpose**

Arterial blood pressure monitoring is vital to clinical decision making. Both the gold standard for monitoring, which is invasive arterial pressure measurement, and the most common modality of blood pressure modality, non-invasive blood pressure monitoring present with limitations that can be resolved with continuous non-invasive arterial pressure (CNAP) monitoring. The aim of this systematic review is to summarize the evidence regarding the concordance between CNAP and invasive arterial pressure monitoring.

**Methods**

Medline and EMBASE will be searched for studies published using commercially available CNAP monitoring systems that reported on a measurement of the invasive or non-invasive arterial pressure that coincided with a measurement of CNAP. Study selection and quality assessment using the Revised Quality Assessment of Diagnostic Accuracy Studies tool (QUADAS-2) was performed independently by two reviewers.The Grading of Recommendations, Assessment, Development and Evaluations (GRADE) approach was used to summarize the strength of the evidence. Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (population limits of agreement) were calculated. 

**Results**

The primary meta-analysis of CNAP versus invasive arterial pressure consisted of `r primary_sbp_results$Comparisons` comparisons from `r primary_sbp_results$Studies` individual studies, with data from `r primary_sbp_results$Participants` participants with `r format(primary_sbp_results$Measurements, scientfic=F)` paired measurements for SBP were included. For DBP `r primary_dbp_results$Comparisons` comparisons from `r primary_dbp_results$Studies` individual studies, with data from `r primary_dbp_results$Participants` participants with `r format(primary_dbp_results$Measurements, scientfic=F)` paired measurements were included. For MAP `r primary_map_results$Comparisons` comparisons from `r primary_map_results$Studies` individual studies, with data from `r primary_map_results$Participants` participants with `r format(primary_map_results$Measurements, scientfic=F)` paired measurements for MAP were included. The pooled estimate for the mean bias was `r primary_sbp_results$bias_mean`mmHg, `r primary_dbp_results$bias_mean`mmHg, and `r primary_map_results$bias_mean`mmHg for SBP, DBP, and MAP, respectively. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary_sbp_results$CI_L_rve` mmHg to `r primary_sbp_results$CI_U_rve`mmHg for SBP, `r primary_dbp_results$CI_L_rve` mmHg to `r primary_dbp_results$CI_U_rve`mmHg for DBP, and  `r primary_map_results$CI_L_rve`mmHg to `r primary_map_results$CI_U_rve`mmHg for MAP. 

**Conclusion** 




Clinical trial number: Not applicable

Keywords: arterial pressure, continuous non-invasive monitoring, critical care 


\pagebreak

## Declarations

Availability of data and material (data transparency): All data used in the meta-analyses is available [here]().


Code availability: All data used in the meta-analyses is available [here](https://github.com/nkamboj06/cnap-review) and archived [here]().

Abbreviated title: 

Funding statement: No funding recieved. 

Conflicts of interest: The authors declare no competing interests.

Author contributions: 
NK: Study design, analysis, wrote manuscript; AC: Study design, revised the manuscript for important intellectual content; KC: Study design, revised the manuscript for important intellectual content; KM: Study design, revised the manuscript for important intellectual content; CC: Study design, revised the manuscript for important intellectual content.

**Glossary of terms:**

GRADE = Grading of Recommendations, Assessment, Development and Evaluations

QUADAS-2 = Revised Quality Assessment of Diagnostic Accuracy Studies tool

LoA = Limits of Agreement

CI = Confidence intervals

SD^2^ = Variance

&tau;^2^ = Tau-squared

\pagebreak

## Introduction

Blood pressure monitoring is a mainstay of hemodynamic monitoring and is a requirement for patients in critical care settings [@Meidert2018Techniques] For patients requiring continuous blood pressure monitoring, invasive arterial pressure monitoring is often used as the standard of care and involves arterial cannulation. [@Meidert2018Techniques] The cannulation of an artery is painful, time-consuming, needs to be done by a trained clinician, and is associated-although very rarely- with complications such as infections, embolism, tissue and nerve damage. [@Meidert2018Techniques] In addition, specialized equipment is required in order to analyse arterial pressure continuously. For these reasons, non-invasive blood pressure (NIBP) monitoring is frequently used. [@Meidert2018Techniques] While there are many benefits of NIBP monitoring, the devices using this method are unable to provide continuous blood pressure measurements and may leave blood pressure fluctuations undetected or may lead to late recognition and delayed correction. For this reason, continuous non-invasive arterial pressure (CNAP) monitoring is fast gaining importance as it addresses the complications associated with the invasive monitoring and the limitations of NIBP monitoring. 
	
CNAP monitoring is based on the arterial applanation tonometry (AAT) and volume clamp. Both techniques enable the arterial waveform and BP values to be obtained continuously. [@Stenglova2017CNAP] AAT is based on the work of Pressman and Newgard, who discovered that a transducer strapped to an artery with a bone underneath, can obtain the arterial pulse wave. [@Meidert2018Techniques] A device that is automated and commercially available that uses this method is the T-Line system (Tensys Medical, San Diego, CA, USA).
Volume clamp is based on the work by Penaz et al. (1976). [@Penaz1976Volumeclamp] and measure blood pressure at the finger using an inflatable cuff combined with a photodiode. [@Meidert2018Techniques] Devices using this technique include Nexfin (BMEYE B.V., Amsterdam, The Netherlands); CNAP (CNSystems, Graz, Austria). CNAP monitoring devices display real-time, continuous arterial pressure waveforms and provide non-invasive beat-to-beat arterial pressure measurement. Numerous studies have investigated the concordance between CNAP measurement and invasive arterial pressure measurement across a variety of clinical populations. Critical appraisal of the quality of these studies followed by synthesis of results in a meta-analysis would aid clinical decision-making regarding the appropriateness of substituting CNAP for blood pressure monitoring in clinical practice.

\pagebreak

## Methods

A systematic review was conducted. The primary comparison for this review was blood pressure measured using CNAP monitoring devices versus blood pressure measured using an invasive device.

### Inclusion criteria

Studies that reported blood pressure measurements using CNAP technique and comparator invasive were included. Due to the potential of overestimation of the intervention performance, case control design studies were excluded. Studies were limited to human subjects admitted on an in-patient clinical healthcare setting (not including operative room). No publication date or language restrictions were applied. Published conference abstracts were included if there was enough information reported to appraise the quality of the study. 

### Data sources and searches

The information extracted included study characteristics (author, year of publication, country, design, sample size, clinical setting, numbers studied, and analyses for each outcome), population characteristics (inclusion and exclusion criteria) and blood pressure measurement characteristics (type of CNAP device and site of invasive measurements). The outcomes that were extracted included the mean bias (eg, accuracy) and variance (eg, SD, precision) in SBP, DBP, and MAP measurement between the invasive and CNAP devices.  We also extracted information about how repeated measurements were handled. In particular we assessed whether studies: (1) analysed each pair of data separately; (2) treated each pair of data as independent; or (3) used either analysis of variance or a random effects model as a way to control for the dependent nature of the repeated measures data.[@myles2007using]

Two reviewers independently completed the risk of bias assessment for the included studies using the revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2).[@whiting2011quadas] Risk of bias for patient selection, conduct of the CNAP measurements, conduct of the comparator invasive measurements, and timing and flow (eg, timing of CNAP and established invasive blood pressure measurements, dropouts) was rated as 'high', 'low' or 'unclear' risk of bias by the reviewers. 

A comprehensive search of multiple databases and international clinical trial registry was conducted to minimize the risk of publication bias. [@glasziou2001systematic] Due to the lack of validated methods on statistical approaches for detecting reporting bias, this assessment was not performed. [begg2005systematic] Simulations have revealed that tests for detecting funnel plot asymmetry will result in publication bias being frequently identified incorrectly. [@deeks2005performance]

The Grading Quality of Evidence and Strength of Recommendation methodology was applied to rate the quality of evidence.[@schunemann2008grading] Evidence was downgraded in accordance with study limitations, inconsistency and imprecision. There were no circumstances in which evidence was downgraded for indirectness as this systematic review only included relevant studies. Although the possibility of publication bias was not excluded, this bias was not formally assessed as it was not considered sufficient enough to reason downgrading the quality of evidence.

### Data synthesis and analysis

The objective for the meta-analysis was to estimate the population limits of agreement between blood pressure measurements from the devices using the CNAP monitoring devices and established invasive monitoring. A framework for meta-analysis of Bland-Altman method comparison studies based on limits of agreement approach was used.[@tipton2017framework]  This method was selected because it mirrors the approach in primary Bland-Altman studies, providing an estimate of the pooled limits of agreement in the population (not just the samples studied). The ‘population LoA’ is wider than those typically reported in meta-analyses of Bland-Altman studies.[@tipton2017framework] Pooled limits of agreement were calculated using $\delta\pm2\sqrt{\sigma2+\tau2}$, where $\delta$ is the average bias across studies, $\sigma2$ is the average within-study variation in differences and $\tau2$ is the variation in bias across studies. 

We estimated $\delta$ and $\sigma2$ using a weighted least-squares model (similar to a random-effects approach) and associated estimations of their SEs were made using robust variance estimation (RVE). As some studies included used repeated-measures designs without accommodating for the correlation between measurements, robust variance estimation was used instead to model-based standard errors. [@hedges2010robust; @tanner2016handling; @tipton2015small] The method-of-moments estimator from DerSimonian & Laird @dersimonian1986meta was used for the $\tau2$ parameter.

The outer 95% confidence intervals for pooled limits of agreement were calculated to determine the measures of uncertainty. If the individual studies did not account for repeated measurements, it was adjusted for in our study by using weights proportional to the number of participants and not the total number of measurements. All the analyses were conducted using the R statistical program.[@R2018] All data and R code used in the meta-analyses is available [here](https://github.com/nkamboj06/cnap-review) and archived [here]().

(NEED TO WRITE THIS PARAGRAPH- IGNORE) Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator invasive blood pressure measurements minus the CNAP device measurement. In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'.[@Boisson_2018; @Dahyot_Fizelier_2017; @Schell_Chaple_2018; @M_kinen_2016] Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups.[@West_2019; @Pesonen_2018; @M_kinen_2016; @Iden_2015] One study used both nasopharyngeal and oropharyngeal temperature as comparators.[@West_2019] Both combined and separate estimates were reported. We used the combined estimate in our primary analysis and the estimate for just the nasopharyngeal temperatures in a subgroup analysis.  One study reported intraoperative, postoperative and overall results for the same participants[@eshraghi2014]. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups. The conventionally cited clinically acceptable agreement between zero-heat-flux and comparator devices is 0.5°C.[@eshraghi2014] It was deemed that outer confidence bounds for 95% limits of agreement between zero-heat-flux and core temperature measurements (termed as ‘population limits of agreement’) outside of these bounds would be clinically unacceptable. 

A sensitivity analysis for the primary comparison (CNAP versus invasive arterial pressure) was performed based on risk of bias. Studies rated as ‘unclear risk of bias’ was treated as ‘high risk’ and ‘high risk of bias’ studies were excluded from analyses. We conducted subgroup analyses for the primary comparison according to the method of CNAP monitoring (Volume Clamp or Arterial Applanation Tonometry (AAT)), type of CNAP monitoring device (Finapres, NCAT, CNAP, Nexfin, T-Line), and measurement site of invasive arterial pressure (radial vs. femoral).

## Results

### Study selection and description

Ninteen studies were included (Figure 1). Twenty studies were conducted in the wrong setting (eg, outpatient, OR) and were excluded. 7 studies were on the wrong population (i.e <18 years of age, subjects were volunteers), and two studies reported the wrong outcome, therefore were excluded.  

A summary of the characteristics of included studies is displayed in Table 1. The Nexfin and T-Line device were both equally evaluated, followed by the CNAP device, Finapres, and NCAT. From the 19 studies, seven used the femoral artery and 12 used the radial artery for invasive measurement.The primary comparison of the CNAP versus invasive arterial pressure measurements consisted of 262, 352 measurements, 785 participants, and 19 comparisons from 18 individual studies of SBP. Primary DBP comparison consisted of 262, 235 measurements, 760 participants, and 18 comparisons from 17 studies. Primary MAP comparison consisted of 162, 080 measurements, 784 participants, and 19 comparisons from 18 individual studies. 

### Primary Comparison

Table 2 presents results of the primary, sensitivity, and all subgroup analyses. The pooled estimate for the mean bias between the CNAP and invasive arterial pressure measurements was -1.91, 1.51, and 1.49 mmHg for SBP, DBP, and MAP, respectively.  Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from – 59.93 to 56.11 mmHg, -216.70 to 219.72 mmHg, and -68.89 to 71.87 for SBP, DBP, and MAP, respectively. Primary SBP comparison consisted of 262, 352 measurements,785 participants, 18 studies. Primary DBP comparison consisted of 262, 235 measurements, 760 participants, and 17 studies. Primary MAP comparison consisted of 162, 080 measurements, 784 participants, and 18 studies. The amount of between-study heterogeneity is displayed graphically in the density plot in Fig.3. 

### Sensitivity Analyses 

### Subgroup Analyses

We conducted three subgroup analyses for the primary comparison according to the method of CNAP monitoring, type of CNAP monitoring device, and measurement site of invasive arterial pressure. In the subset of studies conducted using a device based on the volume clamp method, the mean bias and population limits of agreement was -1.63 (-97.44 to 94.17) for SBP, -1.30 (-320.25 to 317.65) for DBP, and 0.83 (-137.63 to 139.28) for MAP. For the AAT method the mean bias and population limits of agreement was -2.41 (-54.74 to 49.91) for SBP, 5.54 (-31.22 to 42.30) for DBP, and 2.15 (-21.59 to 25.89) for MAP. In the subset of studies conducted based on the type of CNAP monitoring device, the mean and population limits of agreement for Nexfin was -1.09 (-295.59 to 293.42) for SBP, -2.71 (-221.51 to 216.09) for DBP, and -0.41 (-58.24 to 57.42) for MAP. For CNAP Monitoring System it was -3.15 (-155.61 to 149.32) for SBP, -1.28 (-1430.91 to 1428.36) for DBP, and 1.68 (-501 to 504.36) for MAP. For Finapres Medical Systems device it was …..NEED TO DISCUSS WITH AARON. In the subset of studies conducted using the femoral site for invasive arterial pressure measurement, the mean bias and population limits of agreement was -2.98 (-311.89 to 305.93) for SBP, 2.38 (-588.85 to 593.61) for DBP, and 0.09 (-18.71 to 18.89) for MAP. For the radial site it was -0.92 (-70.03 to 68.19) for SBP, 2.09 (-285.69 to 289.87) for DBP, and 1.16 (-80.36 to 82.67) for MAP. 


\pagebreak

## Discussion 

This systematic review showed that a temperature measurement from a 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System device could be as much as about 1°C higher or lower than core temperature. These results may have important implications for practice. It was reassuring that results of our sensitivity analysis restricted to studies assessed to be at low risk of bias using the QUADAS-2 tool were similar. As such, it is vital for healthcare professionals considering using this device to first determine if differences in temperature smaller than this magnitude would be important for the given clinical situation. If so, then it may not be appropriate to substitute the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System in place of a core thermometer. 

Our estimates of the accuracy of measurements from the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System are similar to results from a previous meta-analysis that compared other peripheral thermometers, such as sublingual and temporal artery devices, with core temperature measurements.[@niven2015accuracy] However, this previous meta-analysis used a statistical approach that did not incorporate the magnitude of heterogeneity in results between studies or sampling error. As such, it is possible that the zero-heat-flux thermometer may still be more precise than other peripheral thermometers.

The accuracy of 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System has been evaluated in various intraoperative contexts as well as in the intensive care unit setting. Our subgroup analyses did not indicate that the accuracy of this temperature monitoring device was any more effective in a particular setting and revealed insights to direct future research. There were fewer participants included in the subgroup of studies conducted in the Intensive Care Unit compared to those conducted during surgery. As a result, population limits of agreement for the Intensive Care Unit subgroup were broad. Additional studies are required to increase confidence in the accuracy of 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System device in this setting, where continuous temperature monitoring is often required. Also of note, only two studies included in our systematic review included pediatric patients. Additional studies to evaluate the accuracy of this device in children may be warranted to increase confidence. 

It should be noted that we did not include studies that used the Temple Touch Pro because it is not strictly a zero-heat-flux device. This is a new thermometer that is similar to the zero-heat-flux device in that it is placed cutaneously, but the underlying technology is different.[@evron2017evaluation] Likewise, studies that evaluated the Dräger Tcore device were not included because this review focused specifically on the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System.

Many studies in this review analysed a large number of measurements of temperature with relatively small sample sizes. Importantly, the approach we used for our meta-analysis takes this into account. By using robust variance estimation, weights for pooling estimates in the meta-analysis become proportional to the number of patients, not the total number of measurements.[@tipton2017framework]

### Limitations
 
We did not extract data on adverse events due to temperature monitoring with the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System. The possibility of publication bias cannot be ruled out, although the evidence suggests this may not be as serious of a problem for studies that are not randomized controlled trials.[@begg2005systematic] Our focus for the meta-analysis was on calculating population limits of agreement, which incorporate the variation in bias between studies into the estimates. For this reason, we did not use meta-regression or tests for interaction between subgroups as a way to investigate sources of heterogeneity. It is important to note that this review did not assess the clinical utility of temperature monitoring using the zero-heat-flux device. The evidence from this review should be considered in the context of other information about the reliability and ease of use of this device. For example, the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System permits use of the same temperature monitoring device to be used throughout the whole perioperative care pathway. As such, replacing indirect estimates of core temperature, such as infrared tympanic and temporal devices, may be an advantage of the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System. Indirect estimates of core temperature are not recommended for use in surgical patients, yet are commonly used because of convenience.[@riley2018inadvertent] 

Finally, our initial protocol was focused on the accuracy of zero-heat-flux temperature measurements in general, not the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System specifically. A decision was made after data extraction and analysis was conducted to revise our inclustion criteria and include only studies that reported on the accuracy of the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System for this particular report. This decision was made because of the potential that differences in the technology integrated into zero-heat-flux devices produced by different manufacturers (i.e. 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and the Dräger Tcore) will influence estimates about the accuracy of zero-heat-flux devices overall. Although revising the inclusion criteria must be considered a limitation, this decision permitted a more targeted evaluation of the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and easier interpretation of results.

## Conclusion

Substantial differences between temperature measurements from core and the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System were identified in this meta-analysis. Clinicians should consider the range of uncertainty in the accuracy of the zero-heat-flux thermometer when using this device to inform their decision-making. As such, there may be circumstances where use of this device would not be appropriate because a difference in temperature of 1°C is important to detect for the given clinical situation. 

\pagebreak




# References

<div id="refs"></div>

\pagebreak

# Figure legend
    
Fig. 1 PRISMA Flow Diagram

Fig. 2 Risk of bias assessments for included studies

Fig. 3 Comparisons between invasive and CNAP blood pressure measurement within and across studies


\pagebreak

# Appendix

## Medline search strategy
1.	Blood pressure OR arterial pressure
2.	'non invasive' OR Noninvasive or non-invasive
3.	1 AND 2
4.	Nexfin or Clearsight OR CNAP OR CNAPTM OR Finapres OR Tensys OR T-line OR TL-200 OR TL-300 OR Vasotrac
5.	Penaz OR (Pressman and Newgard) OR Volume Clamp OR Arterial applanation tonometry OR Finger Cuff OR ‘vascular unloading’ OR ‘pulse transit time’
6.	(Continuous OR continued OR continual OR continually OR continuing)
7.	(Beat-to-beat OR real time OR real-time OR simultaneous OR simultaneously)
8.	(Accuracy OR precision OR reliability OR validity OR validation OR standard deviation)
9.	(Bias OR (mean adj1 difference) OR (limi* adj2 agreement) OR (Bland adj1 Altman))
10.	 blood pressure monitors.sh.
11.	6 OR 7
12.	10 OR 11
13.	3 AND 11
14.	5 AND 13
15.	4 OR 12 OR 13 OR 14
16.	8 OR 9
17.	15 AND 16

## EMBASE search strategy
1.	Blood pressure OR arterial pressure
2.	'non invasive' OR Noninvasive or non-invasive
3.	1 AND 2
4.	Nexfin or Clearsight OR CNAP OR CNAPTM OR Finapres OR Tensys OR T-line OR TL-200 OR TL-300 OR Vasotrac
5.	Penaz OR (Pressman and Newgard) OR Volume Clamp OR Arterial applanation tonometry OR Finger Cuff OR ‘vascular unloading’ OR ‘pulse transit time’
6.	(Continuous OR continued OR continual OR continually OR continuing)
7.	(Beat-to-beat OR real time OR real-time OR simultaneous OR simultaneously)
8.	(Accuracy OR precision OR reliability OR validity OR validation OR standard deviation)
9.	(Bias OR (mean adj1 difference) OR (limi* adj2 agreement) OR (Bland adj1 Altman))
10.	 exp blood pressure monitor/
11.	6 OR 7
12.	10 OR 11
13.	3 AND 11
14.	5 AND 13
15.	4 OR 12 OR 13 OR 14
16.	8 OR 9
17.	15 AND 16

<!---BLOCK_MULTICOL_START--->


<!---BLOCK_STOP--->


<!--BLOCK_LANDSCAPE_START--->


## Table 1: Study Characteristics

```{r }
drake::readd(characteristics_table)
```

\pagebreak


### Table 2: Results of meta-analyses

```{r }
drake::readd(results_flextable)
```

<!---BLOCK_LANDSCAPE_STOP--->


