---
title: "Accuracy and Precision of Continuous Non-Invasive Arterial Pressure Monitoring in Critical Care"
subtitle: A systematic review and meta-analysis
bibliography: Extras/bibliography.bib
csl: Extras/sp-basic.csl
author:
  - Navpreet Kamboj MScN:
      institute: [LBFON]
      correspondence: true
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Kelly Metcalfe PhD:
      institute: [LBFON, WCH]
      correspondence: false
  - Charlene Chu PhD:
      institute: [LBFON, KITE]
      correspondence: false
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: false

institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - KITE: KITE-Toronto Rehabilitation Institute, University Health Network, Toronto, Canada
  - WCH: Women’s College Research Institute, Toronto, Canada
  
output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE, fig.retina = 5)

library(drake)
library(bookdown)
library(officedown)

loadd(nexfin_results)
loadd(nexfin_dbp_results)
loadd(nexfin_map_results)
loadd(primary_map_results)
loadd(primary_map_lowrisk_results)
loadd(primary_map_nofunding_results)
loadd(primary_sbp_results)
loadd(primary_sbp_lowrisk_results)
loadd(primary_sbp_nofunding_results)
loadd(primary_dbp_results)
loadd(primary_dbp_lowrisk_results)
loadd(primary_dbp_nofunding_results)
loadd(finapres_map_results)
loadd(finapres_sbp_results)
loadd(finapres_dbp_results)
loadd(ncat_map_results)
loadd(ncat_sbp_results)
loadd(ncat_dbp_results)
loadd(tline_map_results)
loadd(tline_sbp_results)
loadd(tline_dbp_results)
loadd(aat_map_results)
loadd(aat_sbp_results)
loadd(aat_dbp_results)
loadd(vc_map_results)
loadd(vc_sbp_results)
loadd(vc_dbp_results)
loadd(radial_map_results)
loadd(radial_sbp_results)
loadd(radial_dbp_results)
loadd(femoral_map_results)
loadd(femoral_sbp_results)
loadd(femoral_dbp_results)
```

\pagebreak

### Abstract

**Objective**

To summarize the evidence regarding the accuracy of continuous non-invasive arterial pressure measurements in adult critical care patients.

**Research Methodology**

Medline, EMBASE, and CINAHL were searched for studies that included adult critical care patients reporting the agreement between continuous non-invasive and invasive arterial pressure measurements. The studies were selected and assessed for risk of bias using the Revised Quality Assessment of Diagnostic Accuracy Studies tool by two independent reviewers. The Grading of Recommendations, Assessment, Development and Evaluations approach was used. Pooled estimates of the mean bias and limits of agreement with outer 95% confidence intervals (termed population limits of agreement) were calculated.

**Results**

Population limits of agreement for systolic blood pressure were wide, spanning from `r primary_sbp_results$CI_L_rve` mmHg to `r primary_sbp_results$CI_U_rve` mmHg (`r primary_sbp_results$Studies` studies; `r primary_sbp_results$Participants` participants). Accuracy of diastolic blood pressure measurements was highly inconsistent across studies, resulting in imprecise estimates for the population limits of agreement.Population limits of agreement for mean arterial pressure spanned from `r primary_map_results$CI_L_rve` mmHg to `r primary_map_results$CI_U_rve` mmHg (`r primary_map_results$Studies` studies; `r primary_map_results$Participants` participants). The evidence was rated as very low-quality due to very serious concerns about heterogeneity and imprecision. 

**Conclusion**

Substantial differences in blood pressure were identified between measurements taken from continuous non-invasive and invasive monitoring devices. Clinicians should consider this broad range of uncertainty if using these devices to inform clinical decision-making in critical care.

Keywords: arterial pressure, arterial pressure monitoring, continuous non-invasive monitoring, critical care

Clinical Implications: 
•	Continuous non-invasive arterial pressure monitoring is not adequately accurate to substitute for invasive monitoring for adult patients in intensive care as large difference in blood pressure measurements between continuous non-invasive and invasive monitoring were identified.
•	The arterial applanation tonometry technique is more accurate than volume clamp for diastolic blood pressure measurement. 
•	The T-line device produces more accurate systolic, diastolic, and mean arterial pressure measurements than ClearSight

\pagebreak

## Declarations

Availability of data and material (data transparency): All data used in the meta-analyses is available [here](https://github.com/nkamboj06/continuous non-invasive arterial pressure-review).

Code availability: All data used in the meta-analyses is available [here](https://github.com/nkamboj06/continuous non-invasive arterial pressure-review) and archived [here](https://doi.org/10.5281/zenodo.4419814).

Abbreviated title: CNAP Monitoring in Critical Care

Funding statement: This research did not receive any specific grant from funding agencies in the public, commercial, or not-for-profit sectors.

Conflicts of interest: The authors declare no competing interests.

Author contributions: NK: Study design, analysis, wrote manuscript; AC: Study design, revised the manuscript for important intellectual content; KC: Study design, revised the manuscript for important intellectual content; KM: Study design, revised the manuscript for important intellectual content; CC: Study design, revised the manuscript for important intellectual content.

**Glossary of terms:**

GRADE = Grading of Recommendations, Assessment, Development and Evaluations

QUADAS-2 = Revised Quality Assessment of Diagnostic Accuracy Studies 

LoA = Limits of Agreement

CI = Confidence intervals

SD^2^ = Variance

τ^2^ = Tau-squared

\pagebreak

## Introduction

For patients requiring continuous blood pressure monitoring, invasive monitoring is considered the gold standard of care [@kim2014accuracy]. Arterial cannulation is painful, time-consuming, needs to be done by a trained clinician, and is associated with complications such as infections, embolism, tissue and nerve damage [@brzezinski2009radial; @haddad2008can]. For these reasons, intermittent non-invasive blood pressure monitoring is a frequently used alternative [@kim2014accuracy]. However, monitoring blood pressure intermittently increases risk of late recognition and delayed correction of hemodynamic compromise [@chen2012impact; @benes2015continuous]. Continuous non-invasive arterial pressure monitoring allows for continuous blood pressure measurements without the risk of vascular complications and inconvenience associated with arterial cannulation. Devices based on this technique display continuous arterial pressure waveforms in real-time and provide non-invasive beat-to-beat blood pressure measurements [@Meidert2018Techniques].

Continuous non-invasive monitoring of blood pressure can be achieved through the arterial applanation tonometry and volume clamp techniques (also called vascular unloading technique or finger cuff technologies) [@Meidert2018Techniques; @teboul2016less; @saugel2014measurement]. Arterial applanation tonometry is based on the work of Pressman and Newgard, who discovered that a transducer strapped to an artery with a bone underneath, can obtain the arterial pulse wave [@Meidert2018Techniques]. A device that is automated and commercially available that uses this method is the T-line system (Tensys Medical, San Diego, CA, USA). The volume clamp technique is based on the work by Penaz et al.(1976) [@Penaz1976Volumeclamp]. Blood pressure is measured at the finger using an inflatable cuff combined with a photodiode [@Meidert2018Techniques]. Devices using this technique include ClearSight system (Edwards Life-sciences, Irvine, CA, USA [formerly known as Nexfin; BMEye, Amsterdam, The Netherlands); and CNAP® Monitor (CNSystems, Graz, Austria). Numerous validation studies have investigated the concordance between continuous non-invasive and invasive arterial pressure measurement in critical care with contradicting results. The purpose of this study is to summarize the evidence regarding the accuracy of continuous non-invasive arterial pressure measurements among patients admitted to an adult intensive care unit. Knowing the accuracy of this method and of the specific continuous non-invasive blood pressure monitoring devices can help clinicians dictate their use for different patient conditions. Critical appraisal of the quality of evidence from these studies followed by synthesis of results in a meta-analysis would aid clinical decision-making regarding the appropriateness of substituting continuous non-invasive for invasive blood pressure monitoring in the intensive care unit. 

\pagebreak

## Methods

We conducted a systematic review and meta-analysis. The primary comparison for this review was blood pressure measured using a continuous non-invasive versus invasive arterial pressure device in adult patients admitted to a critical care setting.

### Inclusion criteria

### Types of Studies 

Studies that reported blood pressure from a continuous non-invasive arterial pressure device with a concurrent invasive measurement obtained from an arterial cannula were included. Studies that examine the concordance between two measurement instruments tend to use a consecutive series design. However, studies using a randomized controlled trial design to evaluate the clinical utility of continuous non-invasive monitoring can also report on the concordance between continuous non-invasive and invasive blood pressure. For this reason, randomized controlled trials were included. Case control design studies were excluded because of the potential risk of overestimating the intervention performance. Published conference abstracts were included if there was adequate information reported to evaluate the quality of the study. The reference lists of previous systematic reviews on continuous non-invasive blood pressure monitoring were reviewed. Other studies included in online repositories of studies on continuous non-invasive blood pressure monitoring devices on the manufacturer's websites were included if the inclusion criteria were met and there was adequate data to perform the meta-analysis.

### Types of Participants

Studies were limited to human subjects who were 18 years of age or older and received care in a critical care setting.

### Types of Index Tests

Studies that measured arterial pressure using continuous non-invasive with invasive blood pressure monitoring were included. This can include published studies using commercially available monitoring systems including (but not limited to) ClearSight system, CNAP® Monitor, T-line, Finapres to compare arterial pressure with that measured by an invasive monitoring system in clinical settings. Although some of these devices are able to monitor other hemodynamic parameters, such as cardiac output, the focus of this study is blood pressure monitoring only.

### Types of Reference Tests

Studies that reported on a measurement of the arterial pressure using invasive arterial pressure monitoring that coincided with a measurement of the index test were included.

### Types of Outcome Measures

Outcomes extracted were the mean bias (e.g., accuracy) and variance (e.g., SD, precision) for systolic blood pressure, diastolic blood pressure, and mean arterial pressure between the continuous non-invasive blood pressure and comparative devices. 

### Data sources and searches

A comprehensive search of EMBASE, Medline, and CINAHL was conducted. There were no publication date or language restrictions applied. The Cochrane Collaboration recommends these databases to be good sources of diagnostic test accuracy studies [@de2008cochrane]. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used [@mann2012should]. This strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies [@preston2015improving]. An academic librarian was consulted to review the search strategy. Databases were searched to March 26th, 2021. The search strategies for each database are presented in appendix one.

### Data extraction and quality assessment

Two independent reviewers screened all studies using Covidence. Any discrepancies were resolved by a third independent reviewer. Data about the study, population, and blood pressure measurement characteristics (type of continuous non-invasive arterial pressure device and site of invasive measurements) was extracted. Outcomes extracted were the mean bias (e.g., accuracy) and variance (e.g., SD, precision) for systolic blood pressure, diastolic blood pressure, and mean arterial pressure. All included studies either reported results of a Bland-Altman analysis or provided the necessary information to calculate the limits of agreement. Information on how repeated measurements were accounted for in the individual studies was extracted. Specifically, we evaluated if each pair of data in the studies was analysed separately; treated as independent; or whether studies used either random effects model or analysis of variance to adjust for the dependent nature of the repeated measurements [@myles2007using].

The revised Quality Assessment of Diagnostic Accuracy Studies (QUADAS-2) tool was used for the risk of bias assessment [@hadian2006evidence]. This tool was completed by two independent reviewers.  Each study was rated as ‘high’, ‘low’ or ‘unclear’ on the following four domains: (1) risk of bias for patient selection; (2) conduct of the continuous non-invasive arterial pressure measurements; (3) conduct of the comparator invasive measurements; and (4) timing and flow. The QUADAS-2 template is presented in appendix two.

The Grading Quality of Evidence and Strength of Recommendation methodology was used to rate the quality of evidence across studies [@schunemann2008grading]. The reason for downgrading the evidence was because of study limitations, inconsistency, and imprecision. Evidence was not downgraded due to indirectness as relevant studies were only included in this systematic review. The possibility of publication bias was not ruled out. However, we did not formally assess for this bias because this was not determined to be an adequate reason to downgrade the quality of evidence.

### Data synthesis and analysis

We used the framework for meta-analysis of Bland-Altman method comparison studies based on the limits of agreement approach [@tipton2017framework]. An advantage of this method is that it incorporates sampling error into the estimate of the pooled limits of agreement. This ‘population limits of agreement’ is broader than those usually reported in meta-analyses of Bland-Altman studies [@tipton2017framework]. A weighted least-squares model was used to estimate average bias across studies and the average within-study variation in differences, similar to the random-effects approach. Estimations of their standard errors were made using robust variance estimation [@hedges2010robust; @tanner2016handling].Some studies applied repeated-measures designs without adjusting for the correlation between measurements from the same participants. The method-of-moments estimator noted by DerSimonian and Laird [@dersimonian1986meta] was applied for the $\tau2$ parameter.

Currently, there are no minimal standards established to determine the accuracy of continuous non-invasive arterial pressure monitoring systems. Criteria developed by the Association for the Advancement of Medical Instrumentation (AAMI, 2009) for evaluating automatic sphygmanometers has been cited in other research to evaluate the accuracy of continuous non-invasive blood pressure measurement devices [@hohn2013non; @martina2012noninvasive; @ilies2015comparison; @ilies2012investigation; @hahn2012clinical; @fischer2012non]. The acceptable limits of an automatic sphygmanometer as defined by AAMI is accuracy no greater than 5 mmHg and precision to be no greater than 8 mmHg for systolic and diastolic blood pressure  [@sphygmomanometers20092]. We calculated the outer 95% confidence intervals for the pooled limits of agreement to identify the measures of uncertainty. If repeated measurements were not adjusted for in the individual studies, we accounted for this in the meta-analysis. This was done by calculating weights proportional to the number of participants instead of the total number of measurements. 

The results from each individual study were standardized, with bias referring to continuous non-invasive minus the comparator invasive blood pressure monitoring system. For the primary comparison (continuous non-invasive versus invasive arterial pressure monitoring systems), a sensitivity analysis was conducted based on the risk of bias assessment results. Studies rated as having ‘unclear risk of bias’ were considered as having ‘high risk’ and all ‘high risk of bias’ studies were not included in this sensitivity analysis. We conducted subgroup analyses for the primary comparison according to the method of monitoring blood pressure (volume clamp vs. arterial applanation tonometry), and arterial catheter site for invasive arterial pressure measurement (radial vs. femoral). As clinicians would benefit from knowing the accuracy of individual devices, we performed a subgroup analyses for the different continuous non-invasive blood pressure monitoring devices, where more than five studies were available. A statistical test to detect publication bias was not performed because previous literature suggests investigating this bias for diagnostic test studies may be problematic or not useful [@song2002asymmetric]. For instance, using asymmetric funnel plots to detect publication bias in reviews of randomized controlled trials have proven to be seriously misleading for diagnostic studies, and alternatives lead to poor power [@leeflang2008systematic]. We used the R statistical program to perform all analyses [@R2018]. To promote transparency, all data and R code used in the meta-analyses can be accessed [here](https://github.com/nkamboj06/cnap-review) and archived [here](https://doi.org/10.5281/zenodo.4419814).

## Results

### Study selection and description

Eighteen studies were included (Figure \@ref(fig:prisma)), and a summary of the characteristics of these studies is presented in table 1. Twenty-nine studies were excluded after full-text review because they did not meet the inclusion criteria or did not report adequate data for the meta-analysis (see appendix three). The primary meta-analysis for systolic blood pressure consisted of `r format(primary_sbp_results$Measurements, scientfic=F)` measurements, `r primary_sbp_results$Participants` participants, and `r primary_sbp_results$Comparisons` comparisons from `r primary_sbp_results$Studies` individual studies. The primary meta-analysis for diastolic blood pressure consisted of `r format(primary_dbp_results$Measurements, scientfic=F)` measurements, `r primary_dbp_results$Participants` participants, and `r primary_dbp_results$Comparisons` comparisons from `r primary_dbp_results$Studies` individual studies. There were `r format(primary_map_results$Measurements, scientfic=F)` measurements from `r primary_map_results$Participants` participants included in the primary meta-analysis for mean arterial pressure (`r primary_map_results$Comparisons` comparisons; `r primary_map_results$Studies` studies). One study reported results for two groups of participants (atrial fibrillation and sinus rhythm) and each of these groups was considered as an individual ‘comparison’ in the meta-analysis [@berkelmans2018comparing].None of the primary studies mentioned excluding patients with hypertension or hypotension. Also, all studies except two (Novak 1994; Searle, 1993) involved patients who were receiving vasoactive drug such as epinephrine, norepinephrine, dobutamine, or milrinone.

The sensitivity analysis of studies rated as having low risk of bias across all domains included `r primary_sbp_lowrisk_results$Comparisons` comparisons from `r primary_sbp_lowrisk_results$Studies` studies that enrolled `r primary_sbp_lowrisk_results$Participants` participants with `r format(primary_sbp_lowrisk_results$Measurements, scientfic=F)` paired measurements for the systolic blood pressure outcome. Primary diastolic blood pressure comparison of low risk of bias studies included `r primary_dbp_lowrisk_results$Comparisons` comparisons from `r primary_dbp_lowrisk_results$Studies` studies that enrolled `r primary_dbp_lowrisk_results$Participants` participants with `r format(primary_dbp_lowrisk_results$Measurements, scientfic=F)` paired measurements. Primary mean arterial pressure comparison for low risk of bias studies included `r primary_map_lowrisk_results$Comparisons` comparisons from `r primary_map_lowrisk_results$Studies` individual studies of `r primary_map_lowrisk_results$Participants` participants with `r format(primary_dbp_lowrisk_results$Measurements, scientfic=F)` paired measurements.

The modified QUADAS-2 was used to conduct the quality assessment (presented in Figure \@ref(fig:rob)). In the patient selection domain, risk of bias was assessed as low in 16 studies, unclear in 1, and high in 1. In the index test domain, 13 studies were assessed as low risk, 3 as unclear risk, and 2 as high risk. In the reference standard domain, 12 studies were assessed as low risk, 2 as unclear risk, and 4 as high risk. In the flow and timing domains, 11 studies were low risk, 3 unclear risk, and 4 high risk. Overall, 7 studies were assessed as low risk and 11 as high risk. In 7 (39%) studies, the authors declared a receipt of funding from manufacturers, institutional, departmental, or an infrastructure grant.

### Primary meta-analysis

The results of the primary, sensitivity, and subgroup analyses are presented in table 2. The pooled estimate for the mean bias between the continuous non-invasive and invasive arterial pressure measurements was `r primary_sbp_results$bias_mean` mmHg, `r primary_dbp_results$bias_mean` mmHg, and `r primary_map_results$bias_mean` mmHg for systolic, diastolic, and mean arterial pressure, respectively. Population limits of agreement, which account for the between-study heterogeneity and sampling error, were wide, ranging from `r primary_sbp_results$CI_L_rve` mmHg to `r primary_sbp_results$CI_U_rve` mmHg for systolic blood pressure, `r primary_dbp_results$CI_L_rve` mmHg to `r primary_dbp_results$CI_U_rve` mmHg for diastolic blood pressure, and `r primary_map_results$CI_L_rve` mmHg to `r primary_map_results$CI_U_rve` mmHg for mean arterial pressure. The density plots display the amount of between-study heterogeneity (Figure \@ref(fig:combined)). The quality of evidence for the primary meta-analysis was downgraded to very low quality due to concerns about imprecision, inconsistency, and study limitations.

### Sensitivity Analyses

The sensitivity analysis for studies rated as low risk across all domains of the QUADAS-2 revealed wider population limits of agreement than the primary meta-analysis for all outcomes. The mean bias for systolic blood pressure was `r primary_sbp_lowrisk_results$bias_mean` mmHg with population limits of agreement ranging from `r primary_sbp_lowrisk_results$CI_L_rve` mmHg to `r primary_sbp_lowrisk_results$CI_U_rve` mmHg. The mean bias for diastolic blood pressure was `r primary_dbp_lowrisk_results$bias_mean` mmHg with population limits of agreement of `r primary_dbp_lowrisk_results$CI_L_rve` mmHg to `r primary_dbp_lowrisk_results$CI_U_rve` mmHg. The mean bias for mean arterial pressure was `r primary_map_lowrisk_results$bias_mean` mmHg, with population limits of agreement between `r primary_map_lowrisk_results$CI_L_rve` mmHg and `r primary_map_lowrisk_results$CI_U_rve` mmHg. The quality of evidence for this sensitivity analysis was downgraded to very low quality due to concerns about imprecision and inconsistency. A sensitivity analysis not including studies that received any type of funding showed population limits of agreement that were wider than the primary analysis (`r primary_sbp_nofunding_results$CI_L_rve` mmHg to `r primary_sbp_nofunding_results$CI_U_rve` mmHg for systolic blood pressure, `r primary_dbp_nofunding_results$CI_L_rve` mmHg to `r primary_dbp_nofunding_results$CI_U_rve` mmHg for diastolic blood pressure, and `r primary_map_nofunding_results$CI_L_rve` mmHg to `r primary_map_nofunding_results$CI_U_rve` mmHg for mean arterial pressure). The GRADE rating for quality of evidence for this sensitivity analysis was reduced to very low, again, because of concerns about inconsistency, study limitations, and imprecision.

### Subgroup Analyses

Three subgroup analyses were performed based on the method of continuous non-invasive arterial pressure monitoring, types of monitoring device, and the anatomical site (femoral or radial artery) where the arterial catheter was placed for invasive blood pressure measurement. In the subset of studies conducted using a device based on the arterial applanation tonometry, the pooled mean bias (population limits of agreement) for mean arterial pressure was `r aat_map_results$bias_mean` mmHg (`r aat_map_results$CI_L_rve` mmHg to `r aat_map_results$CI_U_rve` mmHg). In the subset of studies conducted using a device based on the volume clamp method, the pooled mean bias (population limits of agreement) for mean arterial pressure was `r vc_map_results$bias_mean` mmHg (`r vc_map_results$CI_L_rve` mmHg to `r vc_map_results$CI_U_rve` mmHg) for mean arterial pressure. The evidence was downgraded to very low quality because of concerns about imprecision, inconsistency, and study limitations.

In the subset of studies conducted based on the monitoring device, the pooled mean bias (population limits of agreement) of the mean arterial pressure for the ClearSight system was `r nexfin_map_results$bias_mean` (`r nexfin_map_results$CI_L_rve` mmHg to `r nexfin_map_results$CI_U_rve` mmHg) for mean arterial pressure. In the subset of studies based on the type of T-line monitoring device, the mean bias (population limits of agreement) was `r tline_map_results$bias_mean` (`r tline_map_results$CI_L_rve` mmHg to `r tline_map_results$CI_U_rve` mmHg) for mean arterial pressure. The GRADE rating for quality of evidence was downgraded to very low due to study limitations, imprecision, and inconsistency.

In the subset of studies conducted using the femoral site for invasive mean arterial pressure measurement, the pooled mean bias (population limits of agreement) was `r femoral_map_results$bias_mean` mmHg (`r femoral_map_results$CI_L_rve` mmHg to `r femoral_map_results$CI_U_rve` mmHg). For the radial site, the estimates for pooled mean bias (population limits of agreement) was `r radial_map_results$bias_mean` mmHg (`r radial_map_results$CI_L_rve` mmHg to `r radial_map_results$CI_U_rve` mmHg) for mean arterial pressure. The quality of evidence was downgraded to very low quality due to concerns about imprecision, inconsistency, and study limitations. The meta-analysis results for systolic and diastolic blood pressure for studies included in the subgroup analysis are identified in table 2.

One study used the femoral artery as the comparator for all 45 participants and completed an additional analysis on 17 of the 45 participants using the radial artery for invasive MAP measurement [@de2014validation]. The results from the radial artery were included in the measurement site sub-group analysis, whereas the results from the femoral artery were reported in the primary analysis. Two studies used a mixture of radial and femoral artery for comparator measurements [@lakhal2016cnap; @hohn2013non]. The authors of these studies reported the combined results therefore these two studies were not included in the arterial catheter site sub-group analysis.
\pagebreak

## Discussion

We found that systolic and mean arterial blood pressure measurements from continuous non-invasive arterial pressure monitoring devices could be as much as about 30-45 mmHg lower or higher than an invasive measurement recorded at the same time. Results for the diastolic blood pressure outcome suggest continuous non-invasive arterial pressure devices produce highly inaccurate measurements for this parameter. Based on these results, it is not appropriate to substitute blood pressure measurements from continuous non-invasive devices in place by those obtained from invasive arterial pressure monitoring in critical care settings when close hemodynamic monitoring is required. The evidence was downgraded to very low quality due to concerns about imprecision, inconsistency and risk of bias.

Pooled estimates for the accuracy of continuous non-invasive arterial pressure devices have been reported in previous meta-analyses [@kim2014accuracy; @saugel2020continuous]. Our estimates for the population limits of agreement should be considered a better representation of the magnitude of discrepancy which should be expected between continuous non-invasive and invasive arterial pressure measurements if these devices were to be used in critical care settings for a number of reasons. First, in our review, we used the meta-analysis of Bland-Altman method comparison studies framework, which is based on limits of agreement approach [@tipton2017framework]. This approach incorporates the magnitude of heterogeneity in results between studies as well as sampling error. Second, we applied repeated measures adjustments for several of the included studies, which did not account for the correlation between measurements taken from the same participants over time in their primary analyses. This adjustment was not undertaken in the prior meta-analyses, which may lead to more favourable estimates for pooled limits of agreement. Third, by restricting the systematic review to the specific population of critical care patients, this systematic review provides direct evidence for this context. Another distinction is that we included both methods of continuous non-invasive arterial pressure monitoring (volume clamp and arterial applanation tonometry) in our meta-analyses and performed a sub-group analysis. A previous meta-analysis did not conduct a subgroup analysis on these two methods of continuous non-invasive blood pressure monitoring [@kim2014accuracy]. A recent meta-analysis only focused on the volume clamp method [@saugel2020continuous]. Our subgroup meta-analyses revealed important insights into the accuracy of specific types of continuous non-invasive arterial pressure devices in critical care.

The certainty of evidence was downgraded due to imprecision arising from the fact that many included studies analysed a large number of blood pressure measurements from only a small number of participants. For example, one study analysed 8700 measurements from just 29 participants [@martina2014noninvasive]. We used robust variance estimation to ensure that the pooled estimates were proportional to the number of patients, not the total number of measurements [@tipton2017framework]. Therefore, studies with much larger sample sizes are needed to improve confidence in the accuracy and precision of continuous non-invasive arterial pressure devices.

One obvious feature of the primary and sensitivity meta-analyses that requires discussion is that population limits of agreement for diastolic blood pressure were far wider than the systolic and mean arterial pressure outcomes. In many cases, the population limits of agreement even exceeded physiologically plausible ranges, suggesting the non-invasive device was highly inaccurate for diastolic blood pressure measurements. The major cause of the extremely wide population limits of agreement was the considerably large amount of heterogeneity across the studies, which can be evidenced by high values for the τ^2^ parameter (reported in Table 2). The more pronounced variation in limits of agreement for diastolic blood pressure compared with systolic and mean arterial pressure is also clearly visible in the density plots presented in Figure 3. Close inspection of the subgroup analyses based on the type of continuous non-invasive device used, does provide some important additional insight. Population limits of agreement for the diastolic blood pressure outcome in the subgroup meta-analysis of studies that used the arterial applanation tonometry approach were not wider than systolic and mean arterial pressure and did not have high values for the τ^2^ parameter, meaning results between studies were not inconsistent. Based on these results, continuous non-invasive blood pressure devices based on arterial applanation technology should be considered more reliable for use in clinical practice and, in particular, they should be given preference over volume clamp devices for any situation where continuous non-invasive monitoring of diastolic blood pressure is required. It is important to note that technical challenges for measuring blood pressure with arterial applanation tonometry have been reported. For instance, optimal positioning of the sensor is key to obtaining reliable measurements and devices based on this method are very sensitive to muscle contraction and movements of the limb where the device is applied [@abreu2020screening; @kizilova2018review].

The sub-group analysis restricted to studies that used devices based on the arterial applanation tonometry technique was comprised primarily of studies that used only one specific device (T-line). From the seven included studies, six used the T-line and one used the NCAT. In contrast, the subgroup meta-analysis of devices based on the volume clamp approach included three different types of devices (Finapres, CNAP® monitor, ClearSight) across the 11 studies. Moreover, different models for the Finapres, ClearSight and CNAP® devices were used across the studies. It seems, though, that variation in the type of volume clamp device does not explain the large variation in estimates for the accuracy of diastolic blood pressure population limits of agreement. Population limits of agreement for diastolic blood pressure were wide in further subgroup meta-analyses restricted to studies that used the ClearSight device. 

### Limitations

Information on adverse events due to monitoring of blood pressure with continuous non-invasive arterial pressure devices was not extracted. Several different types of continuous non-invasive arterial pressure monitoring devices were combined in our primary meta-analysis. These devices are based on different methods, namely the arterial applanation tonometry and volume-clamp. We performed a sub-group analysis of the different methods (arterial applanation tonometry and volume clamp), and of the different devices (ClearSight system and T-line). Different models of each device were not assessed individually. We did not undertake subgroup meta-analysis for specific continuous non-invasive arterial pressure devices where there was less than five number of studies. For example, there was only one study each exploring the Finapres and NCAT device [@novak1994accuracy; @searle1993assessment].

Our aim was to estimate the absolute agreement between the continuous non-invasive and invasive arterial pressure methods. We found that a blood pressure measurement at a specific point in time cannot be considered an accurate reflection of the 'true' (e.g., invasive) blood pressure. As blood pressure continuously changes overtime in response to a number of factors, an alternative way that continuous non-invasive devices could potentially be useful in clinical practice is if the blood pressure trends over time were accurate [@saugel2020continuous; @smolle2015accuracy]. However, we did not evaluate the trending ability because methods to meta-analyses results from individual studies are not available.

Two studies [@hansen2006oscillometric; @steiner2003validation] were excluded because there was insufficient data to calculate the limits of agreement. We did not attempt to contact the authors of these studies to request the original data because both were published more than 15 years ago.

Another limitation is that the possibility of publication bias cannot be excluded. The current meta-analysis accounts for the variation in bias between studies in the population limits of agreement estimates. Therefore, using tests for interaction between subgroups or meta-regression to determine causes of heterogeneity is not suitable.

## Conclusion

Substantial differences between blood pressure measurements from continuous non-invasive and invasive monitoring for blood pressure measurements were identified in this meta-analysis. Systolic and mean arterial blood pressure measurements from continuous non-invasive arterial pressure monitoring devices could be as much as about 30-45 mmHg lower or higher than an invasive measurement recorded at the same time. Overall, continuous non-invasive arterial pressure devices produced highly inaccurate measurements for diastolic blood pressure. However, non-invasive devices that use arterial applanation tonometry technology to measure blood pressure differ from invasive measurements by 30 mmHg lower to 45 mmHg higher range. Population limits of agreement were wide, spanning from `r primary_sbp_results$CI_L_rve` mmHg to `r primary_sbp_results$CI_U_rve`mmHg for systolic blood pressure, `r primary_dbp_results$CI_L_rve` mmHg to `r primary_dbp_results$CI_U_rve` mmHg for diastolic blood pressure, and `r primary_map_results$CI_L_rve` mmHg to `r primary_map_results$CI_U_rve` mmHg for mean arterial pressure. Clinicians should consider this broad range of uncertainty of continuous noninvasive blood pressure monitoring if using these devices to inform clinical decision-making in critical care.

\pagebreak

# References

::: {#refs}
:::

\pagebreak

# Figure legend

Fig. \@ref(fig:prisma) PRISMA Flow Diagram.

Fig. \@ref(fig:rob) Risk of bias assessments for included studies.

Fig. \@ref(fig:combined) Comparisons between continuous non-invasive and invasive blood pressure measurement within and across studies. Blue curves are distributions of the differences between measurements from  continuous non-invasive and invasive arterial blood pressure measurements in individual studies. The red curve is the distribution of the pooled estimate.

\pagebreak

```{r prisma, fig.cap="PRISMA Flow Diagram", fig.height = 6}

drake::readd(prisma)
```

\pagebreak

```{r rob, fig.cap = "Risk of bias assessments for included studies", fig.height = 7.5, fig.width=7}

drake::readd(rob_within)
```

\pagebreak

```{r combined, fig.cap="Comparisons between continuous non-invasive and invasive blood pressure measurement within and across studies. Blue curves are distributions of the differences between measurements from  continuous non-invasive and invasive arterial blood pressure measurements in individual studies. The red curve is the distribution of the pooled estimate.", fig.height = 6}

drake::readd(combined_plot)
```


<!---BLOCK_MULTICOL_START--->

<!---BLOCK_STOP--->

<!--BLOCK_LANDSCAPE_START--->

## Table 1: Study Characteristics

```{r }
drake::readd(characteristics_table)
```

\pagebreak

### Table 2: Results of meta-analyses

```{r }
drake::readd(results_flextable)
```
<!---BLOCK_LANDSCAPE_STOP--->

# Appendix 1

## Medline search strategy

1.  Blood pressure OR arterial pressure
2.  'non invasive' OR Noninvasive or non-invasive
3.  1 AND 2
4.  Nexfin or ClearSight OR CNAP OR CNAPTM OR Finapres OR Tensys OR T-line OR TL-200 OR TL-300 OR Vasotrac
5.  Penaz OR (Pressman and Newgard) OR Volume Clamp OR Arterial applanation tonometry OR Finger Cuff OR 'vascular unloading' OR 'pulse transit time'
6.  (Continuous OR continued OR continual OR continually OR continuing)
7.  (Beat-to-beat OR real time OR real-time OR simultaneous OR simultaneously)
8.  (Accuracy OR precision OR reliability OR validity OR validation OR standard deviation)
9.  (Bias OR (mean adj1 difference) OR (limi\* adj2 agreement) OR (Bland adj1 Altman))
10. blood pressure monitors.sh.
11. 6 OR 7
12. 10 OR 11
13. 3 AND 11
14. 5 AND 13
15. 4 OR 12 OR 13 OR 14
16. 8 OR 9
17. 15 AND 16

## EMBASE search strategy

1.  Blood pressure OR arterial pressure
2.  'non invasive' OR Noninvasive or non-invasive
3.  1 AND 2
4.  Nexfin or ClearSight OR CNAP OR CNAPTM OR Finapres OR Tensys OR T-line OR TL-200 OR TL-300 OR Vasotrac
5.  Penaz OR (Pressman and Newgard) OR Volume Clamp OR Arterial applanation tonometry OR Finger Cuff OR 'vascular unloading' OR 'pulse transit time'
6.  (Continuous OR continued OR continual OR continually OR continuing)
7.  (Beat-to-beat OR real time OR real-time OR simultaneous OR simultaneously)
8.  (Accuracy OR precision OR reliability OR validity OR validation OR standard deviation)
9.  (Bias OR (mean adj1 difference) OR (limi\* adj2 agreement) OR (Bland adj1 Altman))
10. exp blood pressure monitor/
11. 6 OR 7
12. 10 OR 11
13. 3 AND 11
14. 5 AND 13
15. 4 OR 12 OR 13 OR 14
16. 8 OR 9
17. 15 AND 16

\pagebreak

## CINAHL Plus search strategy

1.  Blood pressure OR arterial pressure
2.  'non invasive' OR Noninvasive or non-invasive
3.  1 AND 2
4.  Nexfin or ClearSight OR CNAP OR CNAPTM OR Finapres OR Tensys OR T-line OR TL-200 OR TL-300 OR Vasotrac
5.  Penaz OR (Pressman and Newgard) OR Volume Clamp OR Arterial applanation tonometry OR Finger Cuff OR 'vascular unloading' OR 'pulse transit time'
6.  (Continuous OR continued OR continual OR continually OR continuing)
7.  (Beat-to-beat OR real time OR real-time OR simultaneous OR simultaneously)
8.  (Accuracy OR precision OR reliability OR validity OR validation OR standard deviation)
9.  (Bias OR (mean adj1 difference) OR (limi\* adj2 agreement) OR (Bland adj1 Altman))
10. (MH ‘blood pressure monitor’)
11. 6 OR 7
12. 10 OR 11
13. 3 AND 11
14. 5 AND 13
15. 4 OR 12 OR 13 OR 14
16. 8 OR 9
17. 15 AND 16

\pagebreak

# Appendix 2: Quality Assessment by Using Revised Quality Assessment of Diagnostic Accuracy Studies Tool (QUADAS-2)

Patient selection	
  Was a consecutive or random sample of patients enrolled?	High/ Low / Unclear
	Was a case-control design avoided?	High/ Low / Unclear
	Did the study avoid inappropriate exclusions?	High/ Low / Unclear
	What was the overall risk of bias associated with the selection of participants?	High/ Low / Unclear

Index test
  Were the continuous non-invasive arterial pressure results interpreted without knowledge of the results of     the comparator?	High/ Low / Unclear
	Were the continuous non-invasive arterial pressure device measurements performed accurately? 	High/ Low /     Unclear
	What was the overall risk of bias associated with the conduct or interpretation of continuous non-invasive     arterial pressure results?	High/ Low / Unclear
	
Reference standard
	Were comparator blood pressure results interpreted without knowledge of continuous non-invasive arterial      pressure measurements?	High/ Low / Unclear
	Was the comparator blood pressure device used the same way for all patients?	High/ Low / Unclear
	Were the comparator measures performed accurately?	High/ Low / Unclear
	What was the overall risk of bias associated with the conduct or interpretation of comparator blood          pressure measurements?	High/ Low / Unclear

Flow and Timing
	Was the interval between continuous non-invasive arterial pressure and comparator blood pressure              measurement device the same for all patients and repeated measurements? 	High/ Low / Unclear
	Were all patients included in the analysis?	High/ Low / Unclear
	Were repeated measurements taken into account in the Bland-Altman analysis?	High/ Low / Unclear
	What was the overall risk of bias associated with the participant flow?	High/ Low / Unclear

\pagebreak

# Appendix 3: List of 29 Excluded Studies

Hansen, S., & Staber, M. (2006). Oscillometric blood pressure measurement used for calibration of the arterial tonometry method contributes significantly to error. Eur J Anaesthesiol, 23(9), 781-7. 

Reason: Insufficient or lack of data (only reported the worst and lowest range in SBP error). 

Steiner, L. A, Johnston, A. J, Salvador, R., Czosnyka, M., & Menon, D. K. (2003). Validation of a tonometric noninvasive arterial blood pressure monitor in the intensive care setting. Anaesthesia, 58, 448–54 

Reason: Inadequate data (results reported as bias and 10th and 90th percentile. Standard deviation could not be calculated).

Jagadeesh, A. M., Singh, N. G., & Mahankali, S. (2012). A comparison of a continuous noninvasive arterial pressure (CNAP) monitor with an invasive arterial blood pressure monitor in the cardiac surgical ICU. Annals of Cardiac Anaesthesia, 15(3), 180–4 

Reason: Patients under 18 years of age (inclusion criteria for their study is age >16 years).

Brittain, J. M., Busk, T. M., & Møller, S. (2018). Validation of non-invasive haemodynamic methods in patients with liver disease: the Finometer and the Task Force Monitor. Clin Physiol Funct Imaging, 38(3), 384-389. 

Reason: Wrong setting (not critical care)

Gomez-Angelats, E., Sierra, C., Coca, A., Pare, J. C., & de la Sierra, A. (2004). Lack of association between blood pressure variability and left ventricular hypertrophy in essential hypertension. Medicina Clinica, 123(19), 731-734.

Reason: Wrong setting (not critical care)

Kawahara, M. (1990). Evaluation of the accuracy of non-invasive automatic blood pressure monitors. Anesthesia progress, 37(5), 244-247.

Reason: Wrong setting (operative room- patients having oral surgery under general anesthesia) 

Kugler, J., Rollnik, J., & Schmitz, N. (1997). Retest-reliability and convergent validity of noninvasive blood pressure determination: arm sphygmomanometry vs. Penaz-method. International journal of clinical monitoring and computing, 14(4), 251-254.

Reason: Wrong setting (not critical care)

Lal, S. K., Mihailidou, A. S., Cejnar, M., Henderson, R. J., Jones, M., & Hunyor, S. N. (1993). Continuous, non-invasive volume-clamp blood pressure: determinants of performance. Journal of hypertension, 11(12), 1413-1422.

Reason: Wrong setting (did not specify critical care and sample included 60 volunteers)

Tanaka, S. Nogawa, M. Yamakoshi, T. & Yamakoshi, K. (2007). Accuracy assessment of a noninvasive device for monitoring beat-by-beat blood pressure in the radial artery using the volume-compensation method. IEEE transactions on bio-medical engineering, 54(10), 1892-5.

Reason: Wrong population (sample consisted of nine healthy adult subjects aged from 22 to 48 years)

Nelesen, R. A., & Dimsdale, J. E. (2002). Use of radial arterial tonometric continuous blood pressure measurement in cardiovascular reactivity studies. Blood pressure monitoring, 7(5), 259-63.

Reason: Wrong setting (did not specify critical care, recruited participants in hypertension and sleep apnea studies conducted at the University of California, San Diego)

Jagomagi, K., Raamat, R., & Talts, J. (2001). Effect of altering vasoactivity on the measurement of finger blood pressure. Blood pressure monitoring, 6(1), 33-40

Reason: Wrong population (participants were healthy volunteers). 

Imholz, B. P., van Montfrans, G. A., Settels, J. J., van der Hoeven, G. M., Karemaker, J. M. & Wieling, W. (1988). Continuous non-invasive blood pressure monitoring: reliability of Finapres device during the Valsalva manoeuvre. Cardiovascular research, 22(6), 390-7.

Reason: Wrong setting (not in critical care) 

Hehenkamp, W. J. K., Rang, S., van Goudoever, J., Bos, W. J. W., Wolf, H., & van der Post, J. A. M. (2002). Comparison of Portapres with standard sphygmomanometry in pregnancy. Hypertension in pregnancy, 21(1), 65-76.

Reason: Patients under 18 years of age (age range in this study is 15-43 years).

Gizdulich, P., Prentza, A., & Wesseling, K. H. (1997). Models of brachial to finger pulse wave distortion and pressure decrement. Cardiovascular research, 33(3), 698-705.

Reason: Wrong population (included healthy volunteer)

Eeftinck Schattenkerk, D. W., van Lieshout, J. J., van den Meiracker, A. H., Wesseling, K. R., Blanc, S., Wieling, W., van Montfrans, G. A., Settels, J. J., Wesseling, K. H., & Westerhof, B. E. (2009). Nexfin noninvasive continuous blood pressure validated against Riva-Rocci/Korotkoff. American journal of hypertension, 22(4), 378-83

Reason: Wrong patient population (volunteers) 

Belani, K. G., Buckley, J. J., & Poliac, M. O. (1999). Accuracy of radial artery blood pressure determination with the Vasotrac. Canadian journal of anaesthesia = Journal canadien d'anesthesie, 46(1), 488-96. 

Reason: Wrong patient population (healthy volunteers)

Wagner, J. Y., Prantner, J. S., Meidert, A. S., Hapfelmeier, A., Schmid, R. M., & Saugel, B. (2014). Noninvasive continuous versus intermittent arterial pressure monitoring: evaluation of the vascular unloading technique (CNAP device) in the emergency department. Scandinavian journal of trauma, resuscitation and emergency medicine, 22(101477511), 8

Reason: Wrong setting (Emergency department) 

van Egmond, J., Hasenbos, M., & Crul, J. F. (1985). Invasive v. non-invasive measurement of arterial pressure. Comparison of two automatic methods and simultaneously measured direct intra-arterial pressure. British journal of anaesthesia, 57(4), 434-44.
Reason: Wrong setting (included ICU and operative room) 

van der Does, Y., van Loon, L. M., Alsma, J., Govers, A., Lansdorp, B., Rood, P. P. M., & Schuit, S. C. E. (2013). Non-invasive blood pressure and cardiac index measurements using the Finapres Portapres in an emergency department triage setting. The American journal of emergency medicine, 31(7), 1012-6. 

Reason: Wrong setting (Emergency department) 

Schutte, A. E., Huisman, H. W., van Rooyen, J. M., Malan, N. T., & Schutte, R. (2004). Validation of the Finometer device for measurement of blood pressure in black women. Journal of human hypertension, 18(2), 79-84.

Reason: Wrong setting (not in critical care) 

Pouwels, S., Lascaris, B., Nienhuijs, S. W., Arthur Bouwman, R., & Buise, M. P. (2017). Validation of the Nexfin non-invasive continuous blood pressure monitoring validated against Riva-Rocci/Korotkoff in a bariatric patient population. Journal of clinical anesthesia, 39(an9, 8812166), 89-95.

Reason: Wrong setting (Outpatient clinic)

Parati, G., Casadei, R., Groppelli, A., Di Rienzo, M., & Mancia, G. (1989). Comparison of finger and intra-arterial blood pressure monitoring at rest and during laboratory testing. Hypertension, 13(6), 647-55.

Reason: Wrong setting (not in critical care)

Nowak, R. M., Sen, A., Garcia, A. J., Wilkie, H., Yang, J. J., Nowak, M. R., & Moyer, M. L. (2011). Noninvasive continuous or intermittent blood pressure and heart rate patient monitoring in the ED. The American journal of emergency medicine, 29(7), 782-9.

Reason: Wrong setting (Emergency department)

Idzenga, T., Reesink, K. D., van Swelm, Y., Hansen, H. H. G., Holewijn, S., & de Korte, C. L. (2012). Noninvasive estimation of the blood pressure waveform in the carotid artery using continuous finger blood pressure monitoring. Ultrasound in medicine & biology, 38(11), 1998-2006. 

Reason: Wrong population (included healthy volunteers)

Epstein, R. H., Huffnagle, S., & Bartkowski, R. R. (1991). Comparative accuracies of a finger blood pressure monitor and an oscillometric blood pressure monitor. Journal of clinical monitoring, 7(2), 161-7.

Reason: Wrong setting (operative room)

Broch, O., Bein, B., Gruenewald, M., Carstens, A., Illies, C., Schoneich, F., Steinfath, M., & Renner, J. (2013). A comparison of continuous non-invasive arterial pressure with invasive radial and femoral pressure in patients undergoing cardiac surgery. Minerva anestesiologica, 79(3), 248-56.

Reason: Wrong setting (operative room)

Arfeen, Z. U., Maran, N. J., Simon, E. J., & McClure, J. H. (1996). A comparison of non-invasive methods of blood pressure measurement in normotensive and hypertensive pregnant women. International journal of obstetric anesthesia, 5(3), 168-71.

Reason: Wrong setting (not in critical care; outpatient clinics) 

Allan, P. D., O'Donnell, T., & Tzeng, Y. (2018). Agreement between finger plethysmography- and brachial oscillometry-derived blood pressure measurements. Clinical physiology and functional imaging, 38(3), 439-446. 

Reason: Wrong population (included healthy volunteers; data obtained from the Wellington Cerebral Hemodynamics Database)

Akkermans, J., Diepeveen, M., Ganzevoort, W., van Montfrans, G. A., Westerhof, B. E., & Wolf, H. (2009). Continuous non-invasive blood pressure monitoring, a validation study of Nexfin in a pregnant population. Hypertension in pregnancy, 28(2), 230-42.

Reason: Wrong setting (outpatient clinic) 